---
title: "Interest Rate Shock Model"
author: "Alain LeBel"
date: "20 February 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source('main_GlobalReport.R', echo = FALSE)
source('report_chart_PCGLUF_factor_model_results_over_time.R', echo = FALSE)
source('bond_vs_bmark_ts.R', echo = FALSE)
source('history_of_shock_forecasts.R', echo = FALSE)
```
## What is the _Interest Rate Shock Model_?
<!-- Remember, you can render this in Visual Studio by going:
	rmarkdown::render('preso.Rmd')
at the command line.  In practice, these presos are easier to work on from RStudio. -->
> "All models are wrong but some are useful" --- George Box

Start with multi factor model with only two factors, being the benchmark equity index ($B$) and the bond index ($F$ for fixed income):
$$r_t = \alpha + \beta_B r_{B,t} + \beta_F r_{F,t} + \theta_t$$
<!-- (The $\theta_t$ is a residual term which we assume will always average out to zero.  The $\alpha$ ought to be very close to zero.)  --> 
For every stock, I run a **multilinear regression** using three years of weekly historic data to determine the values of $\alpha$ and the two $\beta$'s. The result is a simple linear model: plug in returns of the two factors to get an estimated return of the stock. 

To build a complete _Interest Rate Shock Model_, we use simpler 1-factor models to estimate how yield changes affect the benchmark and bond returns, then plug those estimates into the above equation.

## The factors and the shock parameter
For the benchmark, use the _MSCI AC World_ total return index.

For the yield, use the _U.S. 10y Treasury Yield_.  When I shock the yield (by increasing it by 100 bp), I actually compute the log return of the latest yield:
$$log(`r round(1 + yield_index[nrow(yield_index),1], 2)`\%) - log(`r round(yield_index[nrow(yield_index),1], 2)`\%) =
`r round(log(1 + yield_index[nrow(yield_index),1]) - log(yield_index[nrow(yield_index),1]), 4)`$$
is the "return of the yield" shock.  I use this to estimate the shocked returns of the benchmark and bond index.

For the bond, use _Citigroup US Broad Investment-Grade Treasury Bond Index_.  It measures "the total rate of return performance for bond markets with a remaining maturity of at least one year" and is "composed of US Treasuries excluding Federal Reserve purchases, inflation-indexed securities and STRIPS" _(?)_.  Northfield tends to use Citigroup fixed-income indices, and for this yield I find the correlation to be very good (see next page).

## How yield affects bond index returns
_Returns of the yield_ are highly correlated to bonds:  saying "The (return of the) yield is $+x\%$" is almost the same as saying "The bond index goes down $`r round(lm(bond ~ yield, asset_bmark_yield_df )$coefficients[2], 3)[1]`x\%$":

```{r bond-yield regression, echo=FALSE, fig.height=5}
show_regression(asset_bmark_yield_df, "bond", "yield")
```
<!-- This chart should look good.  Point out that this is *not* the part of the model we should worry about. -->

## How yield affects benchmark returns
> "The fools are certain and the intelligent full of doubt." --- Bertrand Russell

_Don't forget correlation:_  If the bond changes, then the benchmark will change, too:

```{r benchmark-bond regression, echo=FALSE, fig.height=5}
show_regression(asset_bmark_yield_df, "bmark", "bond")
``` 
<!-- This chart should look bad.  Say:  "This ought to ought to underline the imperfection in the models.  Note that the 'big boy' risk models are no better, drawing 'lines drawn clouds of dots' just like this." -->

## Equities vs fixed income throughout history
Disclaimer:  benchmarks and bonds now have low correlation _(remember:  correlation$^2 = R^2$)_ and are inversely related ($\beta < 0$)

```{r benchmark-bond relationships over time, echo=FALSE, fig.height=6}
ggplot(melted_ts_combo, aes(end_date, value, colour = benchmark)) + geom_line() + ggtitle("Benchmarks vs Bond:  Beta & R-squared (4y monthly rolling)") + facet_wrap(~metric) + theme(legend.position="bottom")
```

## Why 3 years weekly?
We tried all timespans from six months to five years, with daily, weekly and monthly frequencies:

```{r pfolio_rel_estimates_over_times, echo=FALSE, fig.height = 6}
ggplot(df_ts_experiment, aes(months, pfolio_return_rel_bmark, colour = frequency)) + geom_line() + ggtitle("Next pfolio rel return if yield +100 bp:  model result by timespan and frequency")
```

## Now the results:
I compute this model for each stock in the global portfolios (GLUF and Global Pensions).  Then I shock the 10Y U.S. Treasury Yield by +100 bps and calculate the estimated return of your portfolio based on latest weights in the database.  I find:

- PCGLUF estimated return = `r paste(round(100 * df_portfolio_PCGLUF$pfolio_return_delta_shock_unadjusted_by_weighted,2), "%", sep = "")`.  

- benchmark estimated return = `r paste(round(100 * df_portfolio_PCGLUF$bmark_return_delta_shock,2), "%", sep = "")`.  

- Therefore, we estimate PCGLUF would underperform by `r paste(round(100 * (df_portfolio_PCGLUF$pfolio_return_delta_shock_unadjusted_by_weighted - df_portfolio_PCGLUF$bmark_return_delta_shock),2), "%", sep = "")` in this scenario.

Similarly, PCGPEN estimated return is `r paste(round(100 * df_portfolio_PCGPEN$pfolio_return_delta_shock_unadjusted_by_weighted,2), "%", sep = "")`, or `r paste(round(100 * (df_portfolio_PCGPEN$pfolio_return_delta_shock_unadjusted_by_weighted - df_portfolio_PCGPEN$bmark_return_delta_shock),2), "%", sep = "")` relative the MSCI AC World Total Return Index.

## Checks and biases as of Feb 2017
<!-- “Beware of bugs in the above code; I have only proved it correct, not tried it.” --- Donald Knuth -->
I also computed the model for _every stock in the MSCI AC World_:

* **Internal consistency:** Benchmark estimated shocked return is 3.80% when computed in this way.  Compare this with 3.75% shocked return we get from our simpler linear model.

* **Currency bias:** When I restrict my by-stock calculations to US stocks, my total shocked return of the MSCI US stocks is 4.5%.

* **Industry groups:** Banks (+11.4%), Diversified Financials (+9.1%), Energy (+7.1%), Insurance (+6.7%), Automobiles (+6.5%) and Semiconductors (+6.4%) are the industry groups with the highest shocked returns.  Lowest are Utilities (-4.0%), Real Estate (-2.5%), Food/Beverage/Tobacco (-0.4%) and Household/Personal Products (+0.5%)

## How the results changed through history

```{r, echo = FALSE}
ggplot(filtered_melted_df, aes(end_date, value, colour = variable)) + geom_line() + ggtitle(paste(pfolio_code, "How shock forecasts changed over time")) + theme(legend.position="bottom")
```

## How does this compare to Barra?
I also computed the model over the same timeframe which Barra did for us a couple of months ago (October 2013 to October 2016).  I find:

* PCGLUF estimated return is +3.0%, while Barra computed +6.4%

* the benchmark estimated returns +4.2%, while Barra computed +8.0%

* My model says -1.2% underperformance, while Barra say -1.6%

## Why so different from other risk models?
The two main differences with a true risk model (by Northfield, Barra or UBS) are:

1. Most risk models use "exponentially decay weighted" observations.  This means history has a half-life, and the model puts more emphasis on recent data than older data.

2. Other risk models have more factors, including:
	+ Sectors (GICS or otherwise)
	+ Regions or countries
	+ Currencies
	+ Oil prices
	+ The usual quant signals (size, Value/Growth, etc)
	+ Some economic signals (IP, Slope of the Term Structure, etc)
  
