---
title: "Interest Rate Shock Model"
author: "Alain LeBel"
date: "16 February 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source('main_GlobalReport.R', echo = FALSE)
source('report_chart_PCGLUF_factor_model_results_over_time.R', echo = FALSE)
```

## What is the _Interest Rate Shock Model_?
Start with multi factor model with only two factors, being the benchmark equity index ($B$) and the bond index ($F$ for fixed income):
$$r_t = \alpha + \beta_B r_{B,t} + \beta_F r_{F,t} + \theta_t$$
<!-- (The $\theta_t$ is a residual term which we assume will always average out to zero.  The $\alpha$ ought to be very close to zero.)  --> 
For every stock, I run a **multilinear regression** using three years of weekly historic data to determine the values of $\alpha$ and the two $\beta$'s. The result is a simple linear model: plug in returns of the two factors to get an estimated return of the stock. 

To build a complete _Interest Rate Shock Model_, we use simpler 1-factor models to estimate how yield changes affect the two factor returns, and plug those estimates into the above equation.

## The factors and the shock parameter
For the benchmark, I use the _MSCI AC World_ total return index.

For the bond, I use the total return of the _Citigroup US Broad Investment-Grade Bond Index_ (as does Northfield).  

For the yield, I use the _U.S. 10y Treasury Yield_.

When I shock the yield (by increasing it by 100 bp), I actually compute the return of the latest yield:
$$`r round(1 + us10y_yield[nrow(us10y_yield),2], 2)`\% / `r round(us10y_yield[nrow(us10y_yield),2], 2)`\% =
`r (1 + us10y_yield[nrow(us10y_yield),2]) / us10y_yield[nrow(us10y_yield),2]`$$
is the "return of the yield" shock.  I use this to estimate the returns of the benchmark and bond index.


## How yield affect bond index returns
_Returns of the yield_ <!-- (which I compute as _arithmetic returns_) --> are highly correlated to bonds:  saying "The (return of the) yield goes up $x\%$" is almost the same as saying "The bond index goes down $`r round(lm(bond ~ yield, asset_bmark_yield_df)$coefficients[2], 3)[1]`x\%$":

```{r bond-yield regression, echo=FALSE, fig.height=5}
show_regression(asset_bmark_yield_df, "bond", "yield")
```


## How yield affect benchmark returns
_Don't forget correlation:_  If the bond changes, then the benchmark will change, too:

```{r benchmark-bond regression, echo=FALSE, fig.height=5}
show_regression(asset_bmark_yield_df, "bmark", "bond")
``` 


## Why 3 years weekly?
We tried all timespans from six months to five years, with daily, weekly and monthly frequencies:

```{r pfolio_rel_estimates_over_times, echo=FALSE}
ggplot(df_ts_experiment, aes(months, pfolio_return_rel_bmark, colour = frequency)) + geom_line() + ggtitle("Next pfolio rel return if yield +100 bp:  model result by timespan and frequency")
```

## Now the results:

PCGLUF
PCGPEN

## How does this compare to Barra?

How PCGLUF from Oct compares to the numbers Barra.  
