---
title: "Portfolio shocked by US 10y Treasury Yield +100bp"
output: pdf_document
---

# Portfolio is `r pfolio_code`

<!-- \begin{quote} 
_The fools are certain and the intelligent full of doubt._  
\end{quote}
\hfill -Bertrand Russell
-->

```{r, echo=FALSE, warning=FALSE}
#source('./main_GlobalReport.R')  # <-- You really ought to only run this bit once, and can do outside the R Markdown page.
```

My _Interest Rate Shock Model_ is a multi factor model with only two factors:  the benchmark ($B$) and the bond index ($F$ for factor):
$$r_t = \alpha + \beta_B r_{B,t} + \beta_F r_{F,t} + \theta_t$$
Anyone familiar with the CAPM model might recognise this formula if we drop the $F$ term: $r_t = \alpha + \beta_B r_{B,t} + \theta_t$

For every stock, I run a multilinear regression using three years of weekly historic data to determine the values of $\alpha$ and the two $\beta$'s.  (The $\theta_t$ is 
a residual term which we assume will always average out to zero.  The $\alpha$ ought to be very close to zero.)  

The result is a simple linear model where, if we plug in returns of the factors, we get an estimated return of the stock.  Note that the two factors are correlated:  when I
say "I shock the yield (or bond)", my model is also estimating a return to the benchmark and plugging it into the above linear model.

For the bond, I use the total return of the _Citigroup US Broad Investment-Grade Bond Index_ (as does Northfield).  I use the terms _bond_ and _yield_ interchangeably 
because they are so closely correlated that saying "The _(return of the)_ yield goes up $x\%$" is almost the same as saying "The bond index goes 
down $`r round(lm(bond ~ yield, asset_bmark_yield_df)$coefficients[2], 3)[1]`x\%$":


```{r, echo=FALSE}
show_regression(asset_bmark_yield_df, "bond", "yield")
```

# Results

I computed this model for each stock in the `r pfolio_code` portfolio.  Then I shock the 10Y U.S. Treasury Yield by +100 bps and calculate the estimated return of your portfolio based on
latest weights in the database.  I find:

- your portfolio returns `r paste(round(100 * df_portfolio$pfolio_return_delta_shock_unadjusted_by_weighted,2), "%", sep = "")`.  

- the benchmark returns `r paste(round(100 * df_portfolio$bmark_return_delta_shock,2), "%", sep = "")`.  

- Therefore, the portfolio underperforms by `r paste(round(100 * (df_portfolio$pfolio_return_delta_shock_unadjusted_by_weighted - df_portfolio$bmark_return_delta_shock),2), "%", sep = "")`.  


I also computed the model over the same timeframe which Barra did for us a couple of months ago (October 2013 to October 2016).  I find:

- your portfolio returns +3.0%, while Barra computed +6.4%

- the benchmark returns +4.2%, while Barra computed +8.0%

- My model says -1.2% underperformance, while Barra say -1.6%


The two main differences with a true risk model (by Northfield, Barra or UBS) are:

1. Most risk models use "exponentially decay weighted" observations.  This means history has a half-life, and the model puts more emphasis on recent data than older data.

2. Other risk models have more factors, including:
	+ GICS Sectors
	+ Regions (or countries)
	+ All currencies
	+ Oil prices
	+ The usual quant signals such as size, Value/Growth, etc.

